Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-11-02 doc/paper-lcr.tex
minor editing

@@

% Since voids are not included in the halo model approximation, $\kappah$
% is a biased estimator of $\kappa$, but this can be calibrated for using
% the \proceedure of \Sref{sec:model:voids}. Taking the reconstructed
% $\kappah$ for $10^{5}$ lines of sight we form the joint
% distribution, $\pr(\kappa,\kappah)$ and marginalize over $\pr(\kappah)$ to
% give $\pr(\kappa)$. 
%
% PJM: WTF? We just went through all of this in the previous section!
% We need to check that we are doing what we say we are.

@@

\Fref{fig:where} shows the cumulative contribution to $\kappah$ as a function
of projected distance from the line of sight, magnitude and redshift.
We find that most of the convergence comes from halos whose
galaxies are close to the line of sight. 
Over half of the  convergence
typically comes from halos within 30 arcseconds of the line of sight; by 2
arcminutes, this fraction is \todo{XX\%}{Tom?}.
As a function of magnitude, we find
that $\kappah$ is dominated by objects with magnitudes between $i=18$ and
$i=24$. Objects brighter than $i=18$ are either too rare, or too close to the
observer, to make a significant contribution to the convergence. Objects
fainter than $i=24$ are too small to be important, unless they are extremely
close to the line of sight, but we find that including halos fainter than
$i=24$ does not improve the reconstruction.  \Fref{fig:magcut} shows the
scatter on $\kappah-\kappa$ (where $\kappah$ is shifted so that its mean is
zero) as a function of reconstruction magnitude limit. 
%
% Since these objects
% must be extremely close to the line of sight to make a non-zero $\kappah$
% contribution, it is likely that neglecting stellar mass and using a spherical
% NFW prescription are too naive to adequately reconstruct $\kappax$.
%
% PJM: the sentence above does not make sense!
%
We do
not expect a deeper survey to decrease the size of the uncertainty in mapping
$\kappah$ onto $\kappa$. Halos at all redshifts out to the source redshift
(1.4, here) contribute to the convergence, but the largest contribution comes from
halos with $z \sim z_{\rm source}/2$.

\begin{figure}
\includegraphics[width=\columnwidth]{figs/mag_scatter.eps}
\caption[magcut]{The 16, 50 and 84\% confidence intervals on $\kappah$ minus
$\kappa$ as a function of the limiting $i$ band depth of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:magcut}
\end{figure}

@@

While the \MS catalogs do already contain stellar masses for each galaxy, we
find that these are not well behaved at the low mass end. 
% \todo{Tom}{Explain why we re-made the stellar masses!}. 
We also wanted to be able to perform the functional test of
trying to recover the convergence having assumed the correct stellar
mass--halo mass relation. For these reasons, we assign a new stellar mass to
each halo in the \MS catalogs, according to the stellar mass--halo mass
relation of  \citet{BehrooziEtal2010}. From this true  stellar mass we
simulate an observed stellar mass by drawing a sample from 
$\pr(\log{M^{*}_{\rm obs}}|\log{M^{*}_{\rm true}})$ which we take to be a
Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where
a spectroscopic redshift exists stellar masses can be estimated with typical
uncertainties of 0.15 dex \citep{AugerEtal2009}; however with photometric
redshifts stellar mass uncertainties are typically three times as large. We
use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ otherwise. For photometric redshift errors we draw a
redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a Gaussian
of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$
is the halo's true redshift in the \MS catalogue.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Widthvshilberrt.eps}
\caption{Width of the \infered PDF $\pr(\kappah|\data)$ versus the true
$\kappa$ along a line of sight. Blue assumes a spectroscopic redshift for
every halo, whilst green assumes only photometric redshifts. The width of the
\infered $\pr(\kappah|\data)$ increases with $\kappa$.}
%\todo{Tom}{Check that you mean $\kappah$ and not $\kappa$, and potentially change the graph!}}
\label{fig:widthsvsH}
\end{figure}

@@

% PJM: Note that I removed talk of void correction here - surely we should be
% just applying the method we describe and inferring $\kappa$?

@@

While it is good to reconstruct $\kappa$ precisely, it is more important that
$\kappa$ is reconstructed accurately. 
% A precise, inaccurate $\kappa$ estimator
% is worse than useless for cosmology. 
% PJM: steady on.
We test the accuracy of our reconstructed $\pr(\kappa|\data)$ in the following
way: for each line of sight we shift the reconstructed $\pr(\kappa|\data)$ by
the line of sight's true value of $\kappa$ to give
$\pr(\kappa-\kappatrue|\data)$. Since the $\pr(\kappa-\kappatrue|\data)$ for
each line of sight is then an \independant estimator of zero, the PDFs can be
combined according to
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappa$ estimators can
be considered accurate. 
The point at which the modulus of the bias and the width of $\mathcal{P}_N$
are approximately equal gives an estimate of the number of objects which can
be combined before the measurement is limited by systematic error.

@@

We find that selecting the most tightly constrained lines of sight {\it
decreases} the bias -- our reconstruction is most \succesful for the lowest
$\kappa$ lines of sight. With photometry alone, the bias in this sample is
0.0028. With targeted spectroscopic follow-up of the selected lines of sight
there is negligible improvement, since the best constrained lines of sight are
the most empty, and it is the high-$\kappa$ lines of sight that benefit the
most from spectroscopy. Photometry of the field is sufficient to adequately
select the best lines of sight for cosmology -- and doing so greatly increases
both the precision and accuracy of the cosmological constraints per system.
In the context of time delay distance accuracy given in the previous section,
the sample size would have to be 160 or so before this residual systematic
error was equal to the statistical uncertainty.

@@

Time delay lenses are often selected for follow-up based on their images'
brightness (to make monitoring observations less expensive), their images'
separation (to decrease the covariance between the ground-based image
lightcurves), or the fact that they are quadruply-imaging rather than
doubly-imaging, since this yields 3 time delays instead of 1, and a higher
magnification, more informative Einstein Ring system. All three of these
properties favour lenses with high convergence and shear along their lines of
sight. Focusing on the quad selection, we might expect many of these systems
to have significant external shear (since the cross-section for 4 image
production is so sensitive to ellipticity in the lens mass distribution (or
equivalently, external shear due to the lens' environment).

@@

\item Despite our model's simplicity, its reconstructed $\kappah$ values are
good, if biased, tracers of $\kappa$.  We found that with perfect knowledge of
the halo mass and redshift (from the Millennium Simulation catalogs), for a
typical line of sight the calibrated reconstruction gives an unbiased estimate
of $\kappa$ that is a factor of 2 more precise than the global $\pr(\kappa)$.

\item With uncertain halo masses and redshifts, we find that
$\pr(\kappah|\data)$ is still a useful indicator for inferring
$\pr(\kappa|\data)$ from the ensemble of simulated lines of sight, but the
resulting PDFs tend to be much broader than for perfect halo mass and redshift
reconstructions. The long tails of the $\kappah$ distribution can be
suppressed by focusing on the median value as an indicator for $\kappa$
instead.

@@

\item Selecting the highest precision 33\% of the reconstructed lines of sight
results in a sample with low bias, 0.3\% in terms of the time delay distance,
and a statistical uncertainty due to line of sight mass structure that is half
that of the parent sample. The residual systematic error would be equal to the
overall statistical uncertainty in a sample of around 160 lenses.

\item Conversely, selecting lines of sight with high shear ($|\gamma| >
0.03$ (as might be done inadvertently by focusing on four-image or wide
separation lenses) results in a sample with bias of 1\% in time delay
distance. Such a systematic error would become significant in a sample of 12
lenses.

@@

>>>>

% TEC: yes it's valid. It only really ties us to the halo mass function of the simulation. It's probably tied at the same level as the "global P(\kappa) as derived from Stefan's ray-tracing"

@@

% Since voids are not included in the halo model approximation, $\kappah$
% is a biased estimator of $\kappa$, but this can be calibrated for using
% the \proceedure of \Sref{sec:model:voids}. Taking the reconstructed
% $\kappah$ for $10^{5}$ lines of sight we form the joint
% distribution, $\pr(\kappa,\kappah)$ and marginalize over $\pr(\kappah)$ to
% give $\pr(\kappa)$. 
%
% PJM: WTF? We just went through all of this in the previous section!
% We need to check that we are doing what we say we are.
%
% TEC: yes this doesn't need to be here again
%

@@

\Fref{fig:where} shows the cumulative contribution to $\kappah$ as a function
of projected distance from the line of sight, magnitude and redshift.
We find that most of the convergence comes from halos whose
galaxies are close to the line of sight. 
Over half of the  convergence
typically comes from halos within 30 arcseconds of the line of sight; by 2
arcminutes, this fraction is 85\%. The contribution from halos beyond 2 arcminutes is relatively constant with a contribution of 0.008$\pm$0.003. We find that ignoring halos beyond 2 arcminutes has no effect on the precision of the reconstruction; this is shown in Figure \ref{fig:radcut}.
As a function of magnitude, we find
that $\kappah$ is dominated by objects with magnitudes between $i=18$ and
$i=24$. Objects brighter than $i=18$ are either too rare, or too close to the
observer, to make a significant contribution to the convergence. Objects
fainter than $i=24$ are too small to be important, unless they are extremely
close to the line of sight, but we find that including halos fainter than
$i=24$ does not improve the reconstruction.  \Fref{fig:magcut} shows the
scatter on $\kappah-\kappa$ (where $\kappah$ is shifted so that its mean is
zero) as a function of reconstruction magnitude limit.

% Since these objects
% must be extremely close to the line of sight to make a non-zero $\kappah$
% contribution, it is likely that neglecting stellar mass and using a spherical
% NFW prescription are too naive to adequately reconstruct $\kappax$.
%
% PJM: the sentence above does not make sense!
%
We do
not expect a deeper survey to decrease the size of the uncertainty in mapping
$\kappah$ onto $\kappa$. Halos at all redshifts out to the source redshift
(1.4, here) contribute to the convergence, but the largest contribution comes from
halos with $z \sim z_{\rm source}/2$.

\begin{figure}
\includegraphics[width=\columnwidth]{figs/radius_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappa$ as a function of the limiting radius of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:magcut}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{figs/mag_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappa$ as a function of the limiting $i$ band depth of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:magcut}
\end{figure}

@@

While the \MS catalogs do already contain stellar masses for each galaxy, we
find that these are not well behaved at the low mass end due to unreasonably high dark matter stripping from satellite halos. We also wanted to be able to perform the functional test of
trying to recover the convergence having assumed the correct stellar
mass--halo mass relation. For these reasons, we assign a new stellar mass to
each halo in the \MS catalogs, according to the stellar mass--halo mass
relation of \citet{BehrooziEtal2010}. From this true  stellar mass we
simulate an observed stellar mass by drawing a sample from 
$\pr(\log{M^{*}_{\rm obs}}|\log{M^{*}_{\rm true}})$ which we take to be a
Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where
a spectroscopic redshift exists stellar masses can be estimated with typical
uncertainties of 0.15 dex \citep{AugerEtal2009}; however with photometric
redshifts stellar mass uncertainties are typically three times as large. We
use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ otherwise. For photometric redshift errors we draw a
redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a Gaussian
of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$
is the halo's true redshift in the \MS catalogue.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Widthvshilberrt.eps}
\caption{Width of the \infered PDF $\pr(\kappa|\data)$ versus the true
$\kappa$ along a line of sight. Blue assumes a spectroscopic redshift for
every halo, whilst green assumes only photometric redshifts. The width of the
\infered $\pr(\kappah|\data)$ increases with $\kappa$.}

\label{fig:widthsvsH}
\end{figure}

@@

% PJM: Note that I removed talk of void correction here - surely we should be
% just applying the method we describe and inferring $\kappa$?
%
% TEC: yes.

@@

While it is good to reconstruct $\kappa$ precisely, it is more important that
$\kappa$ is reconstructed accurately. 
We test the accuracy of our reconstructed $\pr(\kappa|\data)$ in the following
way: for each line of sight we shift the reconstructed $\pr(\kappa|\data)$ by
the line of sight's true value of $\kappa$ to give
$\pr(\kappa-\kappatrue|\data)$. Since the $\pr(\kappa-\kappatrue|\data)$ for
each line of sight is then an \independant estimator of zero, the PDFs can be
combined according to
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappa$ estimators can
be considered accurate. 
The point at which the modulus of the bias and the width of $\mathcal{P}_N$
are approximately equal gives an estimate of the number of objects which can
be combined before the measurement is limited by systematic error.

@@

We find that selecting the most tightly constrained lines of sight {\it
decreases} the bias -- our reconstruction is most \succesful for the lowest
$\kappa$ lines of sight. With photometry alone, the bias in this sample is
0.002. With targeted spectroscopic follow-up of the selected lines of sight
there is negligible improvement, since the best constrained lines of sight are
the most empty, and it is the high-$\kappa$ lines of sight that benefit the
most from spectroscopy. Photometry of the field is sufficient to adequately
select the best lines of sight for cosmology -- and doing so greatly increases
both the precision and accuracy of the cosmological constraints per system.
In the context of time delay distance accuracy given in the previous section,
the sample size would have to be 30 or so before this residual systematic
error was equal to the statistical uncertainty.

@@

Time delay lenses are often selected for follow-up based on their images'
brightness (to make monitoring observations less expensive), their images'
separation (to decrease the covariance between the ground-based image
lightcurves), or the fact that they are quadruply-imaging rather than
doubly-imaging, since this yields 3 time delays instead of 1, a higher
magnification and a more informative Einstein Ring system. All three of these
properties favour lenses with high convergence and shear along their lines of
sight. Focusing on the quad selection, we might expect many of these systems
to have significant external shear (since the cross-section for 4 image
production is so sensitive to ellipticity in the lens mass distribution (or
equivalently, external shear due to the lens' environment).

@@

\item Despite our model's simplicity, its reconstructed $\kappah$ values are
good tracers of $\kappa$. $\kappah$ is biased due to our ignorance of voids, but we found that this can be calibrated out. We found that with perfect knowledge of
the halo mass and redshift (from the Millennium Simulation catalogs), for a
typical line of sight the calibrated reconstruction gives an unbiased estimate
of $\kappa$ that is a factor of 2 more precise than the global $\pr(\kappa)$.

\item With uncertain halo masses and redshifts, we find that
$\pr(\kappah|\data)$ is still a useful indicator for inferring
$\pr(\kappa|\data)$ from the ensemble of simulated lines of sight, but the
resulting PDFs tend to be much broader than for perfect halo mass and redshift
reconstructions.

@@

\item Selecting from the highest precision 33\% of the reconstructed lines of sight
results in a sample with low bias, 0.2\% in terms of the time delay distance,
and a statistical uncertainty due to line of sight mass structure that is half
that of the parent sample. Neglecting other sources of error, the residual systematic error would be equal to the
overall statistical uncertainty in a sample of around 30 lenses.

\item Conversely, selecting lines of sight with high shear ($|\gamma| >
0.03$ (as might be done inadvertently by focusing on four-image or wide
separation lenses) results in a sample with bias of 1\% in time delay
distance. Such a systematic error would become significant in a sample of $\sim$8 lenses.

@@

% TEC: I liked it! but it's open to discussion.
